---
title: "hw02_gapminder"
author: "Martin"
output: 
  html_document:
    keep_md: true 
    theme: cerulean
    toc: true
---

```{r, warning=FALSE, echo=FALSE, message=FALSE}
library(gapminder)
library(tidyverse)
library(scales)
library(tsibble)
library(DT)
library(knitr)
library(kableExtra)
```

# Exercise 1.1 - Selected Countries in 1970s

-------------------------------------------------

```{r, warning=FALSE, echo=FALSE, message=FALSE}
gapminder %>%
  filter(year >= 1970 & year <= 1979, 
  country == "Canada" | country == "Brazil" | country == "Japan")%>%
  kable(., format = "html", caption = "Gapminder Data for Brazil, 
  Canada, and Japan")%>%
  kable_styling()
```

# Exercise 1.2 - Selected Countries and GDP Per Capita in the 70s

-------------------------------------------------

```{r, warning=FALSE, echo=FALSE, message=FALSE}
gapminder %>%
  filter(year >= 1970 & year <= 1979, 
  country == "Canada" | country == "Brazil" | country == "Japan") %>%
  select (country, gdpPercap) %>%
  kable(., format = "html", caption = "GDP per Capita in the 1970s") %>%
  kable_styling()
```

# Exercise 1.3 - Drop in Life Expectancy

-------------------------------------------------

```{r, warning=FALSE, echo=FALSE, message=FALSE}
gapminder %>%
  group_by(country)%>%
  arrange(year)%>%
  mutate(increaseinlifeExp = lifeExp - lag(lifeExp)) %>%
  filter(increaseinlifeExp < 0) %>%
  arrange(increaseinlifeExp) %>%
  DT::datatable()
```

# Exercise 1.4 - Max GDP Per Capita for Each Country

-------------------------------------------------

```{r, warning=FALSE, echo=FALSE, message=FALSE}
gapminder %>% 
  group_by(country) %>%
  summarize(maxgdpPercap = max(gdpPercap)) %>%
  DT::datatable()
```


# Exercise 1.5 - Canada Life Expectancy vs GDP Per Capita

-------------------------------------------------

```{r, warning=FALSE, echo=FALSE}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot(., aes(lifeExp, gdpPercap))+
  geom_point(color = "purple")+
  ggtitle("Life Expectancy vs GDP Per Capita in Canada")+
  scale_y_log10("GDP per capita (US$)", labels = 
  scales::dollar_format())+
  xlab("Life Expectancy (Years)")+
  theme_bw()
```


# Exercise 2: Exploring a Quantitative Variable (Population)

-------------------------------------------------

```{r, warning=FALSE, echo=FALSE}
gapminder %>%
  group_by(country) %>%
  summarize(min_pop = min(pop), max_pop = max(pop), mean_pop = 
  mean(pop), median_pop= median(pop), sd_pop = sd(pop)) %>%
  DT::datatable()
```

# Exercise 2: Exploring a Categorical Variable (Continent)

-------------------------------------------------

```{r, warning=FALSE, echo=FALSE}
gapminder %>%
  group_by(continent, country) %>%
  summarize(n_countries = n_distinct(country)) %>%
  ggplot(., aes(continent, n_countries))+
  geom_col(fill = "orange")+
  ggtitle("Number of Countries in Each Continent")+
  ylab("Country Count")+
  xlab("Continent")+
  theme_bw()
```

```{r, warning=FALSE, echo=FALSE}
gapminder %>%
  group_by(continent, country) %>%
  summarize(n_countries = n_distinct(country)) %>%
  tally(n_countries) %>%
  kable(., format = "html", caption = "No. of Countries in Each 
  Continent") %>%
  kable_styling()
```


# Exercise 3: Scatterplot Using mtcars Data Set

-------------------------------------------------

```{r, warning=FALSE, echo=FALSE}
mtcars %>%
  ggplot(., aes(disp, mpg))+
  geom_point(color = "red")+
  ggtitle("Displacement vs MPG from mtcars Data Set")+
  xlab("Displacement(cu.in)")+
  ylab("Miles Per Gallon (MPG)")+
  theme(plot.title = element_text(face = "bold"))+
  theme_bw()
```


# Exercise 3: Bar Graph Using mtcars Data Set

-------------------------------------------------

```{r, warning=FALSE, echo=FALSE}
mtcars %>%
  group_by(gear) %>%
  summarize(n = n()) %>%
  ggplot(., aes(gear, n))+
  geom_col(fill = "dark green")+
  ggtitle("Counts of Each Gear Level")+
  xlab("Gear Level")+
  ylab("Count")+
  theme(plot.title = element_text(face = "bold"))+
  theme_bw()
```



# Recycling

-------------------------------------------------

```{r, warning=FALSE}
filter(gapminder, country == c("Rwanda", "Afghanistan")) %>%
  kable(., format = "html", caption = "Gapminder Data for Rwanda and Afghanistan")%>%
  kable_styling()
```

The result produced was not what the author intended. Instead of filtering all the rows with "Rwanda" and "Afghanistan", they only pulled out 12 rows. 6 rows for "Afghanistan" and 6 rows for "Rwanda". This is incorrect because if we look at the year column, we can see that it does not capture all the data. The gapminder data set contains data for each country from the year 1952 until 2007 in intervals of 5 and each country should have 12 observations (code below). We can see from the output of their code that it does not produce 12 rows for both "Afghanistan" and "Rwanda." In addition, it does not produce all the years correctly for both countries. For example, "Afghanistan" is missing the year 1952. Therefore, based on the number of rows printed and the missing data in the year column, the author's code is not outputting what they intended. 

```{r, warning=FALSE}
unique(gapminder$year)
```

```{r, warning=FALSE}
gapminder %>%
  count(country)
```


What is happening here is that since the "==" operator is being used to test equality on vectors of unequal length, the smaller vector which is "Rwanda" and "Afghanistan," is being recycled until it is the same length as the country vector which happens to be about 1704 rows of data (code below).

```{r, warning=FALSE}
nrow(gapminder)
```

I suspect that R is returning all the **odd numbered rows** which happen to be the **even numbered years** for "Rwanda" and all the **even numbered rows** which happen to be the **odd numbered years** for "Afghanistan" which is why it only outputs 6 rows for both countries. A possible explanation for this is perhaps because "Rwanda" is in the "1" position, it is an odd number, and thus R looks for the first odd numbered row that contains "Rwanda." Then going with this thought, "Afghanistan" is in the "2" position and R looks for the first even numbered that contains "Afghanistan."


# My Fix 

----------------------

To fix the code, here is one proposed solution which outputs 24 rows. 12 for "Afghanistan" and 12 for "Rwanda" from 1952 to 2007 in intervals of 5. 
```{r, warning=FALSE}
gapminder %>%
  filter(country == "Rwanda" | country == "Afghanistan") %>%
  DT::datatable()
```

