---
title: "**Gapminder Exploration**"
output: 
  github_document:
    theme: cerulean
    toc: true
    toc_float: true
    keep_md: true
---
##### This is an excerpt of Gapminder data on life expectancy, GDP per capita, and population by country from 1952 to 2007.
```{r, message=FALSE ,warning=FALSE, echo=FALSE}
library(gapminder)
library(tibble)
```

--------------------------------------------------------------------------

### **How many rows and variables does the gapminder data set contain?**
```{r, warning=FALSE, echo=FALSE}
row <- nrow(gapminder)
col <- ncol(gapminder)
print(paste("This dataset has", row, "rows and", col, "columns."))

```
### **What are the names of the variables?**
```{r, warning=FALSE, echo=FALSE}
names(gapminder)
```
### **Which countries does this dataset contain?**
```{r, warning=FALSE, echo=FALSE}
unique(gapminder$country)
```

### **Which continents does this dataset contain?**
```{r, warning=FALSE, echo=FALSE}
unique(gapminder$continent)

```


### **What is the lowest and highest year this dataset contains? What is the mean and median?**
```{r, warning=FALSE, echo=FALSE}
min <- min(gapminder$year)
max <- max(gapminder$year)
median <- median(gapminder$year)
mean <- mean(gapminder$year)
print(paste("The lowest year is", min, "and the highest year is", max, ". The mean is", mean, "and the median is", median, "."))
```


### **What is the lowest and highest life expectancy? What is the mean and median?**
```{r, warning=FALSE, echo=FALSE}
min <- min(gapminder$lifeExp)
max <- max(gapminder$lifeExp)
median <- median(gapminder$lifeExp)
mean <- mean(gapminder$lifeExp)
print(paste("The lowest life expectancy is", min, "and the highest life expectancy is", max, "years. The mean is", mean, "and the median is", median, "."))

```

### **What is the lowest and highest population? What is the mean and median?**
```{r, warning=FALSE, echo=FALSE}
min <- min(gapminder$pop)
max <- max(gapminder$pop)
median <- median(gapminder$pop)
mean <- mean(gapminder$pop)
print(paste("The lowest population is", min, "and the highest population is", max, ". The mean is", mean, "and the median is", median,"."))

```


### **What is the lowest and highest GDP per capita? What is the mean and median?**
```{r, warning=FALSE, echo=FALSE}
min <- min(gapminder$gdpPercap)
max <- max(gapminder$gdpPercap)
median <- median(gapminder$gdpPercap)
mean <- mean(gapminder$gdpPercap)
print(paste("The lowest GDP per capita is", min, "and the highest GDP per capita is", max, ". The mean is", mean, "and the median is", median, "."))

```

### **Summary of the gapminder data set**
```{r, warning=FALSE, echo=FALSE}
summary(gapminder)

```

### **Conclusion**
This concludes my exploration of the gapminder data set.

