---
title: "Gapminder Exploration"
author: "Martin Keung"
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
    incremental: true
    
  
---

```{r, warning=FALSE, echo=FALSE}
library(gapminder)

```



## **This presentation will cover the following points:**


- How many rows and variables does the gapminder data set contain?
- What are the names of the variables?
- What is the earliest and latest year this dataset contains? What is the mean and median?
- What is the lowest and highest life expectancy? What is the mean and median?
- What is the smallest and largest population? What is the mean and median?
- What is the lowest and highest GDP per capita? What is the mean and median?
- Summary of the Gapminder data set 

## **How many rows and variables does the gapminder data set contain?**
```{r, warning=FALSE, echo=FALSE}
row <- nrow(gapminder)
col <- ncol(gapminder)
print(paste("This dataset has", row, "rows and", col, "variables."))

```

## **What are the names of the variables?**
```{r, warning=FALSE, echo=FALSE}
names(gapminder)
```


## **What is the earliest and latest year this dataset contains? What is the mean and median?**
```{r, warning=FALSE, echo=FALSE}
min <- min(gapminder$year)
max <- max(gapminder$year)
median <- median(gapminder$year)
mean <- mean(gapminder$year)
print(paste("The earliest year is", min, "and the latest year is", max, ". The mean is", mean, "and the median is", median, "."))

```

## **What is the lowest and highest life expectancy? What is the mean and median?**
```{r, warning=FALSE, echo=FALSE}
min <- min(gapminder$lifeExp)
max <- max(gapminder$lifeExp)
median <- median(gapminder$lifeExp)
mean <- mean(gapminder$lifeExp)
print(paste("The lowest life expectancy is", min, "and the highest life expectancy is", max, "years.")) 
print(paste("The mean is", mean, "and the median is", median, "."))

```

## **What is the smallest and largest population? What is the mean and median?**
```{r, warning=FALSE, echo=FALSE}
min <- min(gapminder$pop)
max <- max(gapminder$pop)
median <- median(gapminder$pop)
mean <- mean(gapminder$pop)
print(paste("The smallest population is", min, "and the largest population is", max, ".")) 
print(paste("The mean is", mean, "and the median is", median,"."))

```


## **What is the lowest and highest GDP per capita? What is the mean and median?**
```{r, warning=FALSE, echo=FALSE}
min <- min(gapminder$gdpPercap)
max <- max(gapminder$gdpPercap)
median <- median(gapminder$gdpPercap)
mean <- mean(gapminder$gdpPercap)
print(paste("The lowest GDP per capita is", min, "and the highest GDP per capita is", max,".")) 
print(paste("The mean is", mean, "and the median is", median, "."))

```

## **Summary of the Gapminder data set**
```{r, warning=FALSE, echo=FALSE}
summary(gapminder)

```

## **Ending Remarks**
- This concludes my analysis of the Gapminder data set.
- The only functions in this presentations used were the ones taught from the class meetings 2 and 3.