---
title: "Gapminder Exploration"
author: "Martin Keung"
date: "15/09/2019"
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r}
library(gapminder)
```

```{r, message=FALSE ,warning=FALSE, echo=FALSE}
library(gapminder)
library(tibble)
```

## **The following presentation will cover the following points:**
- How many rows and variables does the gapminder data set contain?
- What are the names of the variables?
- What is the lowest and highest life expectancy? What is the mean and median?
- What is the lowest and highest year this dataset contains? What is the mean and median?
- What is the lowest and highest population? What is the mean and median?
- What is the lowest and highest GDP per capita? What is the mean and median?

## **How many rows and variables does the gapminder data set contain?**
```{r, warning=FALSE, echo=FALSE}
row <- nrow(gapminder)
col <- ncol(gapminder)
print(paste("This dataset has", row, "rows and", col, "columns."))

```

## **What are the names of the variables?**
```{r, warning=FALSE, echo=FALSE}
names(gapminder)
```


## **What is the lowest and highest year this dataset contains? What is the mean and median?**
```{r, warning=FALSE, echo=FALSE}
min <- min(gapminder$year)
max <- max(gapminder$year)
median <- median(gapminder$year)
mean <- mean(gapminder$year)
print(paste("The lowest year is", min, "and the highest year is", max, ". The mean is", mean, "and the median is", median, "."))
```

## **What is the lowest and highest life expectancy? What is the mean and median?**
```{r, warning=FALSE, echo=FALSE}
min <- min(gapminder$lifeExp)
max <- max(gapminder$lifeExp)
median <- median(gapminder$lifeExp)
mean <- mean(gapminder$lifeExp)
print(paste("The lowest life expectancy is", min, "and the highest life expectancy is", max, "years.")) 
print(paste("The mean is", mean, "and the median is", median, "."))

```

## **What is the lowest and highest population? What is the mean and median?**
```{r, warning=FALSE, echo=FALSE}
min <- min(gapminder$pop)
max <- max(gapminder$pop)
median <- median(gapminder$pop)
mean <- mean(gapminder$pop)
print(paste("The lowest population is", min, "and the highest population is", max, ".")) 
print(paste("The mean is", mean, "and the median is", median,"."))

```


## **What is the lowest and highest GDP per capita? What is the mean and median?**
```{r, warning=FALSE, echo=FALSE}
min <- min(gapminder$gdpPercap)
max <- max(gapminder$gdpPercap)
median <- median(gapminder$gdpPercap)
mean <- mean(gapminder$gdpPercap)
print(paste("The lowest GDP per capita is", min, "and the highest GDP per capita is", max,".")) 
print(paste("The mean is", mean, "and the median is", median, x))

```

## **Summary of the gapminder data set**
```{r, warning=FALSE, echo=FALSE}
summary(gapminder)
```

## **Conclusion**
This concludes my analysis of the gapminder data set.