---
title: "**Gapminder Exploration**"
output: 
  github_document:
    theme: cerulean
    toc: true
    keep_md: true
---
### **Introduction**

-----------------------

##### I performed an exploratory data analysis on the Gapminder data set using the functions I learned from class meetings 2 and 3. I also set echo = FALSE for all the code chunks to prevent the code, but not the results from appearing in the finished file. Finally, I tried using the paste function to concatenate the output of the functions with some supporting text. 

```{r, message=FALSE ,warning=FALSE, echo=FALSE}
library(gapminder)

```

--------------------------------------------------------------------------

### **How many rows and variables does this data set contain?**
```{r, message=FALSE, warning=FALSE, echo=FALSE} 
row <- nrow(gapminder)
col <- ncol(gapminder)
print(paste("This dataset has", row, "rows and", col, "variables."))
```
### **What are the names of the variables?**
```{r, message=FALSE, warning=FALSE, echo=FALSE}
names(gapminder)
```
### **Which countries does this data set contain?**
```{r, message=FALSE, warning=FALSE, echo=FALSE}
unique(gapminder$country)
```

### **Which continents does this data set contain?**
```{r, message=FALSE, warning=FALSE, echo=FALSE}
unique(gapminder$continent)
```


### **What is the smallest and largest year this data set contains? What is the mean and median?**
```{r, message=FALSE, warning=FALSE, echo=FALSE}
min <- min(gapminder$year)
max <- max(gapminder$year)
median <- median(gapminder$year)
mean <- mean(gapminder$year)
print(paste("The smallest year is", min, "and the largest year is", max, ". The mean is", mean, "and the median is", median, "."))
```


### **What is the lowest and highest life expectancy? What is the mean and median?**
```{r, message=FALSE, warning=FALSE, echo=FALSE}
min <- min(gapminder$lifeExp)
max <- max(gapminder$lifeExp)
median <- median(gapminder$lifeExp)
mean <- mean(gapminder$lifeExp)
print(paste("The lowest life expectancy is", min, "and the highest life expectancy is", max, "years. The mean is", mean, "and the median is", median, "."))

```

### **What is the smallest and largest population? What is the mean and median?**
```{r, message=FALSE, warning=FALSE, echo=FALSE}
min <- min(gapminder$pop)
max <- max(gapminder$pop)
median <- median(gapminder$pop)
mean <- mean(gapminder$pop)
print(paste("The smallest population is", min, "and the largest population is", max, ". The mean is", mean, "and the median is", median, "."))

```


### **What is the lowest and highest GDP per capita? What is the mean and median?**
```{r, message=FALSE, warning=FALSE, echo=FALSE}
min <- min(gapminder$gdpPercap)
max <- max(gapminder$gdpPercap)
median <- median(gapminder$gdpPercap)
mean <- mean(gapminder$gdpPercap)
print(paste("The lowest GDP per capita is", min, "and the highest GDP per capita is", max, ". The mean is", mean, "and the median is", median, "."))


```

### **Summary of the Gapminder data set**
```{r, warning=FALSE, echo=FALSE}
summary(gapminder)

```

### **Ending Remarks**
This concludes my exploratory data analysis of the Gapminder data set.

